I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-17 Paper/Paper_noProof.tex
update paper

\begin{abstract}
  Deep Neural Networks (DNNs) verification is now a mature research field, with many methodologies and tools to verify formaly  that they are correct, with an annual competition to compare them, etc. Formally, the question is, given a DNN and a property to check, does the property hold over a set of inputs of the DNN. This allows e.g. to check local robustness around an input $I$, by checking that the i-th output neuron has the largest weight among all output neurons uniformly over a neighbourhood around $I$.
  In the most recent years, the focus has been on combining several of these efficient techniques (branch and bound, multi neuron encoding, MILP encoding, ...) to optimize speed/accuracy trade-offs. While some relatively large DNNs (tens of thousands of neurons) can be checked very efficiently with modern verifiers, some DNNs are still challenging to address with efficient algorithms, even small ReLU DNNs with hundreds of neurons, showing the need for new methodologies.

  In this paper, we analyze efficient algorithms to verify ReLU DNNs, based on abstractions (DeepPoly, Linear Programing, PRIMA, different version of Crown), and uncover the main reason for the loss of accuracy, namely {\em compensations}. Intuitively, a compensation happens when there are 2 paths between the same pair of neurons, one path with positive and one path with negative (product of) weights, which therefore (partly) compensate each other. It is however hard to compute exactly by how much is the compensation due to the ReLU activation functions. Based on this finding, we propose a novel methodology to obtain interesting trade-offs in terms of speed/accuracy, using many queries with few 'open' ReLU nodes (for which both linear modes are considered), rather than the usual few queries with potentially many 'open' ReLU nodes. These promising results open up many different applications of the concept of compensating paths in the field of verification of DNNs.
\end{abstract}

@@

While there is one optimal linear function for the upper bound $g_i(x_i)= \beta \frac{x_i-\alpha}{\beta-\alpha}$ for ReLU nodes, there are two choices for the lower bound: $f_i(x_i) = 0$ or $f_i(x_i)=x_i$.
In the DeepPoly algorithm, the choice is made by considering the range of an unstable neuron 
$x_i=[\alpha,\beta]$ with $\alpha < 0 < \beta$. 
If $|\alpha|\geq |\beta|$ then $0$ is used, while if $|\beta|>|\alpha|$ then $x_i$ (the identity function) is used. Let us call $\overline{\mbox{DeepPoly}}$ the Deeppoly algorithm such that 
$f_i(x_i) = 0$ is always used, and $f_i(x_i) = x_i$ is never used. 
It has a main advantage, as it subsmumes the box abstraction.
Indeed, taking an interval $x_i \in [-0.2,5]$, it gives $ReLU(x_i) \in [0,5]$, 
while the normal DeepPoly algorithm would only allow to conclude that $ReLU(x_i) \in [-0.2,5]$.

@@

MILP instances can be linearly relaxed into LP over-abstraction, where variables which were integers in 
${0,1}$ are relaxed to be real in interval $[0,1]$, using otherwise the same encoding. 
Solving LP instances being Polynomial time, much more efficient solution could be found, but at the price of a lack of accuracy (e.g. with not as thight bounds). This is called the {\em LP abstraction}.

As far as we know, no formal result states the relationship between the LP approximation and DeepPoly.
We prove now that the LP abstraction is strictly more accurate tha3 the DeepPoly abstraction, as it actually corresponds to considering in DeepPoly both linear functions lower bounding the ReLU function, $ReLU(y) \geq 0$ and $ReLU(y) \geq y$, while DeepPoly considers exactly one.

@@

It remains to prove that the 2 upper bound contraints of 1) LP corresponds to the 2) DeepPoly linear upper bound. For that, we study the upper bound function in the variable $a \in  [0,1]$, with $y \in [\alpha,\beta]$ fixed. We have $\hat{y}$ in 1) upper bounded by $max_{a \in [0,1]} (min(\beta \cdot a, y - \alpha (1-a)))$, and this bound can be reached. 
The function $min(\beta \cdot a, y - \alpha (1-a))$ reaches its maximum for 
$\beta \cdot a = y - \alpha (1-a)$, that is 
$(\beta - \alpha) a = y - \alpha$, i.e. 
$$ a = \frac{y - \alpha}{\beta-\alpha}$$

@@

Intuitively, the target state receives the weight from the source state from the different paths from the source to the target. As the weights of the paths is of opposite sign, some of these weights will cancel out, by an amount up to $min(|weight(\pi')|,|weight(\pi')|)$. Further, no only do they cancel $min(|weight(\pi')|,|weight(\pi')|)$ out, but they also do not accumulate, in the same way as $0= |1+(-1)| \leq |1|+|-1|=2$ from an inacurate analysis. The amount of compensation can however be reduced by the ReLU functions, because of clipping by the ReLU function. For instance, consider that the inputs $x_1,x_2$ lies in $[-1,-0]$. Then the weight at $\hat{x}_3$ will be always 0, no matter the input, and there will be no compensation to $weight(\pi')$.

@@

Hence, compensation is {\em one} factor of inacurracy. We prove now that it is actually the {\em only} reason for inacurracy, in the following meaning.
If there is no compensating path in the DNN, then any method/approximation at least as accurate as the {\em Box} abstraction, e.g. $\overline{\text{DeepPoly}}$ or {\em LP}, computes the exact upper and lower bounds for every node, that is:

@@

The sketch of proof can be found in Section \ref{sec.proofs}. 
Notice that standard {\em DeepPoly} does not satisfies the following theorem (it does not necessarily refine {\em Box}), and therefore we are not sure PRIMA or $\beta$-CROWN who refine DeepPoly do, though it would be easy to fix by using $\overline{\text{DeepPoly}}$. Anyway, in practice, it is very unlikely to have a network without any compensating path, it is more of a theoretical results showin that compensation is the reason for inacuracy.

@@

Hence $\val_{\vx^\star}(z)=\beta$ and Hence $\val_{\vx^\sharp}(z)=\alpha$,
for $[\alpha, \beta]$ the bounds computed by the Box abstraction. We conclude Theorem \ref{th1} by invocking continuity.

@@

Concerning the Theorem \ref{th2}, consider an output node $z$.
Let $[\alpha(z),\beta(z)]$ be the bound computed by $\MILP_{X_z}$.
We will show the existence of $\vx^\sharp,\vx^\star$
such that $\val_{\vx^\sharp}(z)=\alpha$ and $\val_{\vx^\star}(z)=\beta$,
and we will get the proof of Theorem 2 by a continuity argument.
For that, we partition the input neurons from layer $0$ into four groups:
\begin{enumerate}
	\item $A_{zero}$ is the set of input neurones from where all paths with target $z$ have weight $0$.
	\item $A_{pos}$ is the set of input neurons from where all paths with target $z$ have weight $\geq 0$.
	\item $A_{neg}$ is the set of input neurons from where all paths with target $z$ have weight $\leq 0$.
	\item $A_{open}$ is the set of remaining input neurons.
\end{enumerate}

@@

Using Theorem \ref{th2}, one could thus obtain accurate bounds by running MILP$_{X}$ inductively on every node layer by layer, obtaining accurate bounds $[\alpha,\beta]$ that can be used to compute the next layer accurate bound. In practice, such an algorithm would however be inefficient, as $X_n$ comprises most of the nodes from layer up to $k-1$ for $n$ on layer $k$. Instead, an interesting trade off between speed and accuracy is to set a threshold on the the strength of compensation that are meaningful, and consider that paths below this treshold are not compensating strongly enough to deserve an integer variable (there are cases where the MILP algorithm is exponential in the number of integer variables), calling $Z_n$ the resulting set of nodes on {\em strongly compensating} paths. In practice, one could take the $K$ nodes covering the strongest compensating pair of paths to obtain set $Z_n$, for $K$ reasonably small. 
The pseudo-code of the algorithm is provided in Algorithm \ref{algo1}.

@@

Our experimental evaluations have 3 goals:
\begin{enumerate}
  \item Evaluate experimentally how accurate is $\MILP_Z$ when using compensating strength 
 to choose the set $Z$ of ReLU nodes encoded exactly using an integer variable, compared with randomly chosing the set $Z$, to make sure the choice is relevant. Also compare different choices for $K$ the size of $Z$ and how to rank nodes from different layers,
  \item Evaluate on benchmarks the percentage of certification that can be obtained
 using the compensation idea, and compare with competing efficient tools,
  \item Evaluate the runtime of compensate($K$).
\end{enumerate}

@@

\subsection{Chosing nodes based on Compensation strength}

The first group of experiments we run, tests the accuracy of chosing a set $Z$ of $K$ nodes using compensation strength, compared with randomly chosing the same number $K$ of nodes, to understand if the choice is meaningful or any set $Z$ with the same size will result in a similar accuracy.

To mesure the accuracy, we measure the uncertainty of all nodes in a layer:
the uncertainty of a node is the range between its computed lower and upper bound. 
We then average the uncerainty among all the nodes of the layer.
Formally, for a node $n$ with bounds $[\alpha,\beta]$, its uncertainty $unc(a) = \beta - \alpha$, and the average uncertainty of a layer $l$ is $\dfrac{\sum_{a\in l} unc(a)}{|l|}.$

@@

We report in Table \ref{tab:example0} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly. The range of uncertainty created by inacurate computations is up to $1.17=2.22-1.05$, $2.22$ being the accuracy provided by LP, and $1.05$ by an exact MILP encoding of all the unstable ReLU nodes. Selecting $30$ out of the $61$ unstable ReLU nodes, the uncertainty created by inacurracy is reduced by $80\%$, while a random choice of nodes only leads up to roughly halving this number. We can deduce that selecting nodes based on compensating strength helps to select important ReLU nodes for the accuracy.

@@

We now turn to evaluating chosing ReLU nodes in two consecutive layers, focusing on the uncertainty of nodes in the second layers, wrt ReLU nodes in the first and second layer.
The bounds for nodes of the second layers are computed exactly. 
Let $d$ be a neuron of the third layer.
We keep the previous evaluation for ranking nodes in the previous (second) layer. 
Once the set $Z$ of selected nodes has sufficiently many nodes $c,c'$ in the second layer, adding $b,b'$ to $Z$ allows to take into account accurately the pairs of paths of the form $(a b c d, a b' c' d)$ for all $\{c,c'\} \in Z$. We thus define 
$comp(b,b')=\sum_a \sum_{c,c' \in Z} \val_{\vx}(a) \min(weight(abcd),weight(ab'c'd))$ and the associated $comp(b)$, and select iteratively nodes $b$ (first layer) or $c$ (second layer) by selecting the heaviest $comp(b)$ or $comp(c)$.

We report in Table \ref{tab:example1} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly from layers $1$ and $2$. 
We compare choosing nodes in the second layer only with choosing nodes in both the first and second layer.
Choosing ReLU nodes in the previous layer (layer $2$) only is never a good choice. 
When picking in both layer $1$ and $2$, choosing $50$ out of $116$ unstable ReLU nodes allows to reduce by $90\%$ the average uncertainty due to inacurate computations ($0.059$ vs $0.574$), while it only reduces the inacuracy by $65\%$ when choosing nodes randomly. Notice that this is direct inacuracy, applying this layer after layer will even further reduce inacuracy created by accumulation of inacuraccies.

Notice that it is likely that selecting ReLU nodes up to 3 layers earlier (for nodes in layer $4+$) could further improve the accuracy, but the computational trade off is unclear and we thus didnot experiment.

@@

In terms of efficient verifiers we compare with, we report:
\begin{itemize}
	\item DeepPoly \cite{deeppoly}/ CROWN \cite{crown}. We report the runtime from our own implementation used in Compensate($K$). It shows that our purely Python implementation is not optimal ($10$-$100$ times slower than the implementation in $\alpha$-$\beta$-CROWN, ERAN and PRIMA). Our aim is indeed not to have the most optimized tool, but to evaluate the idea around compensation.
	\item PRIMA, the 2022 version, with results reported in \cite{prima} and \cite{crown}, as it is the closest in practice with our algorithm, computing accurate bounds for all neurons (as we do), although with a different technique: they focus on explicit correlations between neurons, while we consider the correlations indirectly through a better network abstraction. They use GPU while we do not. When computing the whole correlated space, 
	parallelization is difficult.
	Notice that on these MLP networks, PRIMA resort to refined bounds on the first few layers (3 or 4) using exact MILP encoding of all the nodes (which is easily parallelized - 16 nodes considered in parallel in PRIMA). This is doable on the first few layers as the number of integer variable is limited (but stil consequent). By comparison, we run MILP on all the layers, but with a bounded number of integer variables to control the runtime and to scale to all the layers.
	\item $\beta$-CROWN, with results as reported in \cite{crown}, as it is the most accurate of the efficient tool to date. Its main idea is to implement a very efficient Branch and Bound (BaB) verification tool, subsuming \cite{BaB} on such DNNs, and heavily using GPU implementation (while we do not). Performance of pure BaB on these hard DNNs is however impaired by the large number of branchs to consider. Similarly as PRIMA, $\beta$-CROWN resorts to refining the bounds of the first few layers using an exact MILP encoding, also parallelized with 16 nodes in parallel, before running their BaB tool. We also test with the latest versions of $\alpha$-$\beta$-CROWN (October 2023)
	and of PRIMA (2022) on the MLP $6 \times 500$ (results not reported in \cite{crown,prima}).

	\item We do not report results from kPoly \cite{kpoly}, OptC2V \cite{optC2V}, 
	SDPFO \cite{SDPFI} which do not present favorable trade offs compared with $\beta$-CROWN. Results can be found in \cite{crown} and are performed in the same conditions so can be compared.
	\item Finally, we run several version of Compensate($K$). First, we have a fixed choice for $K$ ($K=60$ for fast and $K=200$ for slow), with a fall back to a lower $K$ ($K=50$ for fast and $K=60$ for slow) if the bounds found is too far from the solution found (MIPGap $>0.5$ in Gurobi). We run (our implementation of) DeepPoly first, then the fast version, before running slow on the remaining Images.
	Further, we also experiment with an {\em adaptative} implementation, which lowers number $K$ adaptively between layers to adapt to the growing number $N$ of linear variables, which has an impact on the overall $\MILP(N,K)$ runtime. This is particularly useful for deeper network where $N$ is very different between the first and the last layer. Again, we run first an aggressive version of fast, before going to a more concervative slow version to treat the remaining images. All versions compute bounds for 20 neurons of the same layer in parallel, no GPU used.
\end{itemize}

In Table \ref{tab:example}, we report the accuracy and runtime of these different methods.
DeepPoly/CROWN is by far the fastest, but also the most inacurate, certifying robustness for less than $30\%$ of the images, whereas there are at least $82\%$ of the images for which no attacked is found (attacks computed by \cite{attack} in $\beta$-CROWN).
This is because these DNNs, altough of reasonable size, are hard to verify (they are trained in a natural way, with high compensation strength, see Section \ref{Sec.comp}).

Compared with PRIMA (using refinement of the first few layers by exact MILP), the overall method is quite similar, looking at all the nodes one by one from the begining of the network till the end. Our method compare favorably, both in terms of time and accuracy: our fastest setting is always faster than PRIMA, sometimes by a large margin ($>3$ times faster on $8 \times 100$ while verifying $10\%$ more images), and also more accurate (except our fast adaptative method for $5 \times 200$ which is quite pointless on this shallow DNN - the fixed method is both faster and more accurate than PRIMA), sometimes by a large margin (by $14 \%$  for $5 \times 100$ while being $30\%$ faster). Further, unlike PRIMA, we do not resort to a GPU, and our purely Python implementation is not optimal. This is very promising for this idea of compensation, which indirectly treats dependencies between the variables, which PRIMA represents explicitly.

@@

Compare with $\beta$-CROWN (also using a refinement on the first few layers using exact MILP), 
the picture is more balanced, because the fundamentals of both algorithms are different. BaB 
focuses on the output neuron, computing bounds for it, and refining the different ReLU as necessary. 
Instead, we consider every node one by one from the input layer till the final layer, with the same accuracy. Hence our algorithm will spend a lot of time on nodes which are not necessary for the verification. Therefore, our naive implementation is not as efficient as the highly optimized
$\beta$-CROWN: our implementation is 2 to 5 times slower than $\beta$-CROWN, while we are not using GPU, unlike $\beta$-CROWN. Things are more interesting on the accuracy front: on shallower DNNs ($5 \times 100$ and $5 \times 100$ with 5 hidden layers), the refinement with MILP treats all but the 2 last layers, 
leaving BaB with few ReLU nodes to branch on, and $\beta$-CROWN (refined BaB) is slightly more accuate than using compensation: we are within $1.6\%$ of the accuracy. On deeper DNNs ($8 \times 100$ and $8 \times 100$ with 8 hidden layers), the refinement with MILP cannot treat the 4 last layers, and BaB has more ReLU to branch on. Therefore, Compensate($K$) is more accurate than $\beta$-CROWN, 
closing the gap with the upper bound from $20\%$ to $16.2\%$ ($8 \times 100$)
and from $17.6\%$ to $12\%$ ($8 \times 200$). This is very promising for the compensation idea, which could be used in many different ways than Compensate($K$) (see next Section).

More importantly, $\beta$-Crown, while efficient, often faces a (complexity) wall, when there number of branches becomes too extreme for BaB to work: either BaB succeds fast, or it will not suceed even with very long time out (the reason why Refined BaB is more accurate on these networks). We verify that on a larger DNN learnt in a natural way, namely MLP $6 \times 500$, see Table \ref{tab:example3}.
We tested PureBaB with two time out, from the standard 30s to an extreme 2000s, to match the average run time of our adaptative implementation (fast mode). Only $3.5\%$ more of images could be ceritifed by extending the TO from 30s to 2000s, certifying $20\%$ less images than our prototype in the same time.
Notice that the refined BaB mode of $\beta$-Crown (using MILP for the first few layers) failed to run on this DNN (NoImplementation error code), and we were unable to fix the problem, which is not really surprising as running exact MILP with so many integer variables is not likely to be efficient, as 
confirmed while running PRIMA.

>>>>

\begin{abstract}
  Deep Neural Networks (DNNs) verification is now a mature research field, with many methodologies and tools to verify formally  that they are correct, with an annual competition to compare them, etc. Formally, the question is, given a DNN and a property to check, does the property hold over a set of inputs of the DNN. This allows e.g. to check local robustness around an input $I$, by checking that the i-th output neuron has the largest weight among all output neurons uniformly over a neighbourhood around $I$.
  In the most recent years, the focus has been on combining several of these efficient techniques (branch and bound, multi neuron encoding, MILP encoding, ...) to optimize speed/accuracy trade-offs. While some relatively large DNNs (tens of thousands of neurons) can be checked very efficiently with modern verifiers, some DNNs are still challenging to address with efficient algorithms, even small ReLU DNNs with hundreds of neurons, showing the need for new methodologies.

  In this paper, we analyze efficient algorithms to verify ReLU DNNs, based on abstractions (DeepPoly, Linear Programming, PRIMA, different version of Crown), and uncover the main reason for the loss of accuracy, namely {\em compensations}. Intuitively, a compensation happens when there are 2 paths between the same pair of neurons, one path with positive and one path with negative (product of) weights, which therefore (partly) compensate each other. It is however hard to compute exactly by how much is the compensation due to the ReLU activation functions. Based on this finding, we propose a novel methodology to obtain interesting trade-offs in terms of speed/accuracy, using many queries with few 'open' ReLU nodes (for which both linear modes are considered), rather than the usual few queries with potentially many 'open' ReLU nodes. These promising results open up many different applications of the concept of compensating paths in the field of verification of DNNs.
\end{abstract}

@@

While there is one optimal linear function for the upper bound $g_i(x_i)= \beta \frac{x_i-\alpha}{\beta-\alpha}$ for ReLU nodes, there are two choices for the lower bound: $f_i(x_i) = 0$ or $f_i(x_i)=x_i$.
In the DeepPoly algorithm, the choice is made by considering the range of an unstable neuron 
$x_i=[\alpha,\beta]$ with $\alpha < 0 < \beta$. 
If $|\alpha|\geq |\beta|$ then $0$ is used, while if $|\beta|>|\alpha|$ then $x_i$ (the identity function) is used. Let us call $\overline{\mbox{DeepPoly}}$ the DeepPoly algorithm such that 
$f_i(x_i) = 0$ is always used, and $f_i(x_i) = x_i$ is never used. 
It has a main advantage, as it subsumes the box abstraction.
Indeed, taking an interval $x_i \in [-0.2,5]$, it gives $ReLU(x_i) \in [0,5]$, 
while the normal DeepPoly algorithm would only allow to conclude that $ReLU(x_i) \in [-0.2,5]$.

@@

MILP instances can be linearly relaxed into LP over-abstraction, where variables which were integers in 
${0,1}$ are relaxed to be real in interval $[0,1]$, using otherwise the same encoding. 
Solving LP instances being Polynomial time, much more efficient solution could be found, but at the price of a lack of accuracy (e.g. with not as tight bounds). This is called the {\em LP abstraction}.

As far as we know, no formal result states the relationship between the LP approximation and DeepPoly.
We prove now that the LP abstraction is strictly more accurate that the DeepPoly abstraction, as it actually corresponds to considering in DeepPoly both linear functions lower bounding the ReLU function, $ReLU(y) \geq 0$ and $ReLU(y) \geq y$, while DeepPoly considers exactly one.

@@

It remains to prove that the 2 upper bound constraints of 1) LP corresponds to the 2) DeepPoly linear upper bound. For that, we study the upper bound function in the variable $a \in  [0,1]$, with $y \in [\alpha,\beta]$ fixed. We have $\hat{y}$ in 1) upper bounded by $max_{a \in [0,1]} (min(\beta \cdot a, y - \alpha (1-a)))$, and this bound can be reached. 
The function $min(\beta \cdot a, y - \alpha (1-a))$ reaches its maximum for 
$\beta \cdot a = y - \alpha (1-a)$, that is 
$(\beta - \alpha) a = y - \alpha$, i.e. 
$$ a = \frac{y - \alpha}{\beta-\alpha}$$

@@

Intuitively, the target state receives the weight from the source state from the different paths from the source to the target. As the weights of the paths is of opposite sign, some of these weights will cancel out, by an amount up to $min(|weight(\pi')|,|weight(\pi')|)$. Further, no only do they cancel $min(|weight(\pi')|,|weight(\pi')|)$ out, but they also do not accumulate, in the same way as $0= |1+(-1)| \leq |1|+|-1|=2$ from an inaccurate analysis. The amount of compensation can however be reduced by the ReLU functions, because of clipping by the ReLU function. For instance, consider that the inputs $x_1,x_2$ lies in $[-1,-0]$. Then the weight at $\hat{x}_3$ will be always 0, no matter the input, and there will be no compensation to $weight(\pi')$.

@@

Hence, compensation is {\em one} factor of inaccuracy. We prove now that it is actually the {\em only} reason for inaccuracy, in the following meaning.
If there is no compensating path in the DNN, then any method/approximation at least as accurate as the {\em Box} abstraction, e.g. $\overline{\text{DeepPoly}}$ or {\em LP}, computes the exact upper and lower bounds for every node, that is:

@@

The sketch of proof can be found in Section \ref{sec.proofs}. 
Notice that standard {\em DeepPoly} does not satisfies the following theorem (it does not necessarily refine {\em Box}), and therefore we are not sure PRIMA or $\beta$-CROWN who refine DeepPoly do, though it would be easy to fix by using $\overline{\text{DeepPoly}}$. Anyway, in practice, it is very unlikely to have a network without any compensating path, it is more of a theoretical results show in that compensation is the reason for inaccuracy.

@@

Hence $\val_{\vx^\star}(z)=\beta$ and Hence $\val_{\vx^\sharp}(z)=\alpha$,
for $[\alpha, \beta]$ the bounds computed by the Box abstraction. We conclude Theorem \ref{th1} by invoking continuity.

@@

Concerning the Theorem \ref{th2}, consider an output node $z$.
Let $[\alpha(z),\beta(z)]$ be the bound computed by $\MILP_{X_z}$.
We will show the existence of $\vx^\sharp,\vx^\star$
such that $\val_{\vx^\sharp}(z)=\alpha$ and $\val_{\vx^\star}(z)=\beta$,
and we will get the proof of Theorem 2 by a continuity argument.
For that, we partition the input neurons from layer $0$ into four groups:
\begin{enumerate}
	\item $A_{zero}$ is the set of input neurons from where all paths with target $z$ have weight $0$.
	\item $A_{pos}$ is the set of input neurons from where all paths with target $z$ have weight $\geq 0$.
	\item $A_{neg}$ is the set of input neurons from where all paths with target $z$ have weight $\leq 0$.
	\item $A_{open}$ is the set of remaining input neurons.
\end{enumerate}

@@

Using Theorem \ref{th2}, one could thus obtain accurate bounds by running MILP$_{X}$ inductively on every node layer by layer, obtaining accurate bounds $[\alpha,\beta]$ that can be used to compute the next layer accurate bound. In practice, such an algorithm would however be inefficient, as $X_n$ comprises most of the nodes from layer up to $k-1$ for $n$ on layer $k$. Instead, an interesting trade off between speed and accuracy is to set a threshold on the the strength of compensation that are meaningful, and consider that paths below this threshold are not compensating strongly enough to deserve an integer variable (there are cases where the MILP algorithm is exponential in the number of integer variables), calling $Z_n$ the resulting set of nodes on {\em strongly compensating} paths. In practice, one could take the $K$ nodes covering the strongest compensating pair of paths to obtain set $Z_n$, for $K$ reasonably small. 
The pseudo-code of the algorithm is provided in Algorithm \ref{algo1}.

@@

Our experimental evaluations have 3 goals:
\begin{enumerate}
  \item Evaluate experimentally how accurate is $\MILP_Z$ when using compensating strength 
 to choose the set $Z$ of ReLU nodes encoded exactly using an integer variable, compared with randomly choosing the set $Z$, to make sure the choice is relevant. Also compare different choices for $K$ the size of $Z$ and how to rank nodes from different layers,
  \item Evaluate on benchmarks the percentage of certification that can be obtained
 using the compensation idea, and compare with competing efficient tools,
  \item Evaluate the runtime of compensate($K$).
\end{enumerate}

@@

\subsection{Choosing nodes based on Compensation strength}

The first group of experiments we run, tests the accuracy of choosing a set $Z$ of $K$ nodes using compensation strength, compared with randomly choosing the same number $K$ of nodes, to understand if the choice is meaningful or any set $Z$ with the same size will result in a similar accuracy.

To measure the accuracy, we measure the uncertainty of all nodes in a layer:
the uncertainty of a node is the range between its computed lower and upper bound. 
We then average the uncertainty among all the nodes of the layer.
Formally, for a node $n$ with bounds $[\alpha,\beta]$, its uncertainty $unc(a) = \beta - \alpha$, and the average uncertainty of a layer $l$ is $\dfrac{\sum_{a\in l} unc(a)}{|l|}.$

@@

We report in Table \ref{tab:example0} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly. The range of uncertainty created by inaccurate computations is up to $1.17=2.22-1.05$, $2.22$ being the accuracy provided by LP, and $1.05$ by an exact MILP encoding of all the unstable ReLU nodes. Selecting $30$ out of the $61$ unstable ReLU nodes, the uncertainty created by inaccuracy is reduced by $80\%$, while a random choice of nodes only leads up to roughly halving this number. We can deduce that selecting nodes based on compensating strength helps to select important ReLU nodes for the accuracy.

@@

We now turn to evaluating choosing ReLU nodes in two consecutive layers, focusing on the uncertainty of nodes in the second layers, wrt ReLU nodes in the first and second layer.
The bounds for nodes of the second layers are computed exactly. 
Let $d$ be a neuron of the third layer.
We keep the previous evaluation for ranking nodes in the previous (second) layer. 
Once the set $Z$ of selected nodes has sufficiently many nodes $c,c'$ in the second layer, adding $b,b'$ to $Z$ allows to take into account accurately the pairs of paths of the form $(a b c d, a b' c' d)$ for all $\{c,c'\} \in Z$. We thus define 
$comp(b,b')=\sum_a \sum_{c,c' \in Z} \val_{\vx}(a) \min(weight(abcd),weight(ab'c'd))$ and the associated $comp(b)$, and select iteratively nodes $b$ (first layer) or $c$ (second layer) by selecting the heaviest $comp(b)$ or $comp(c)$.

We report in Table \ref{tab:example1} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly from layers $1$ and $2$. 
We compare choosing nodes in the second layer only with choosing nodes in both the first and second layer.
Choosing ReLU nodes in the previous layer (layer $2$) only is never a good choice. 
When picking in both layer $1$ and $2$, choosing $50$ out of $116$ unstable ReLU nodes allows to reduce by $90\%$ the average uncertainty due to inaccurate computations ($0.059$ vs $0.574$), while it only reduces the inaccuracy by $65\%$ when choosing nodes randomly. Notice that this is direct inaccuracy, applying this layer after layer will even further reduce inaccuracy created by accumulation of inaccuracies.

Notice that it is likely that selecting ReLU nodes up to 3 layers earlier (for nodes in layer $4+$) could further improve the accuracy, but the computational trade off is unclear and we thus did not experiment.

@@

In terms of efficient verifiers we compare with, we report:
\begin{itemize}
	\item DeepPoly \cite{deeppoly}/ CROWN \cite{crown}. We report the runtime from our own implementation used in Compensate($K$). It shows that our purely Python implementation is not optimal ($10$-$100$ times slower than the implementation in $\alpha$-$\beta$-CROWN, ERAN and PRIMA). Our aim is indeed not to have the most optimized tool, but to evaluate the idea around compensation.
	\item PRIMA, the 2022 version, with results reported in \cite{prima} and \cite{crown}, as it is the closest in practice with our algorithm, computing accurate bounds for all neurons (as we do), although with a different technique: they focus on explicit correlations between neurons, while we consider the correlations indirectly through a better network abstraction. They use GPU while we do not. When computing the whole correlated space, 
	parallelization is difficult.
	Notice that on these MLP networks, PRIMA resort to refined bounds on the first few layers (3 or 4) using exact MILP encoding of all the nodes (which is easily parallelized - 16 nodes considered in parallel in PRIMA). This is doable on the first few layers as the number of integer variable is limited (but still consequent). By comparison, we run MILP on all the layers, but with a bounded number of integer variables to control the runtime and to scale to all the layers.
	\item $\beta$-CROWN, with results as reported in \cite{crown}, as it is the most accurate of the efficient tool to date. Its main idea is to implement a very efficient Branch and Bound (BaB) verification tool, subsuming \cite{BaB} on such DNNs, and heavily using GPU implementation (while we do not). Performance of pure BaB on these hard DNNs is however impaired by the large number of branches to consider. Similarly as PRIMA, $\beta$-CROWN resorts to refining the bounds of the first few layers using an exact MILP encoding, also parallelized with 16 nodes in parallel, before running their BaB tool. We also test with the latest versions of $\alpha$-$\beta$-CROWN (October 2023)
	and of PRIMA (2022) on the MLP $6 \times 500$ (results not reported in \cite{crown,prima}).

	\item We do not report results from kPoly \cite{kpoly}, OptC2V \cite{optC2V}, 
	SDPFO \cite{SDPFI} which do not present favorable trade offs compared with $\beta$-CROWN. Results can be found in \cite{crown} and are performed in the same conditions so can be compared.
	\item Finally, we run several version of Compensate($K$). First, we have a fixed choice for $K$ ($K=60$ for fast and $K=200$ for slow), with a fall back to a lower $K$ ($K=50$ for fast and $K=60$ for slow) if the bounds found is too far from the solution found (MIPGap $>0.5$ in Gurobi). We run (our implementation of) DeepPoly first, then the fast version, before running slow on the remaining Images.
	Further, we also experiment with an {\em adaptative} implementation, which lowers number $K$ adaptively between layers to adapt to the growing number $N$ of linear variables, which has an impact on the overall $\MILP(N,K)$ runtime. This is particularly useful for deeper network where $N$ is very different between the first and the last layer. Again, we run first an aggressive version of fast, before going to a more conservative slow version to treat the remaining images. All versions compute bounds for 20 neurons of the same layer in parallel, no GPU used.
\end{itemize}

In Table \ref{tab:example}, we report the accuracy and runtime of these different methods.
DeepPoly/CROWN is by far the fastest, but also the most inaccurate, certifying robustness for less than $30\%$ of the images, whereas there are at least $82\%$ of the images for which no attacked is found (attacks computed by \cite{attack} in $\beta$-CROWN).
This is because these DNNs, although of reasonable size, are hard to verify (they are trained in a natural way, with high compensation strength, see Section \ref{Sec.comp}).

Compared with PRIMA (using refinement of the first few layers by exact MILP), the overall method is quite similar, looking at all the nodes one by one from the beginning of the network till the end. Our method compare favorably, both in terms of time and accuracy: our fastest setting is always faster than PRIMA, sometimes by a large margin ($>3$ times faster on $8 \times 100$ while verifying $10\%$ more images), and also more accurate (except our fast adaptative method for $5 \times 200$ which is quite pointless on this shallow DNN - the fixed method is both faster and more accurate than PRIMA), sometimes by a large margin (by $14 \%$  for $5 \times 100$ while being $30\%$ faster). Further, unlike PRIMA, we do not resort to a GPU, and our purely Python implementation is not optimal. This is very promising for this idea of compensation, which indirectly treats dependencies between the variables, which PRIMA represents explicitly.

@@

Compare with $\beta$-CROWN (also using a refinement on the first few layers using exact MILP), 
the picture is more balanced, because the fundamentals of both algorithms are different. BaB 
focuses on the output neuron, computing bounds for it, and refining the different ReLU as necessary. 
Instead, we consider every node one by one from the input layer till the final layer, with the same accuracy. Hence our algorithm will spend a lot of time on nodes which are not necessary for the verification. Therefore, our naive implementation is not as efficient as the highly optimized
$\beta$-CROWN: our implementation is 2 to 5 times slower than $\beta$-CROWN, while we are not using GPU, unlike $\beta$-CROWN. Things are more interesting on the accuracy front: on shallower DNNs ($5 \times 100$ and $5 \times 100$ with 5 hidden layers), the refinement with MILP treats all but the 2 last layers, 
leaving BaB with few ReLU nodes to branch on, and $\beta$-CROWN (refined BaB) is slightly more accurate than using compensation: we are within $1.6\%$ of the accuracy. On deeper DNNs ($8 \times 100$ and $8 \times 100$ with 8 hidden layers), the refinement with MILP cannot treat the 4 last layers, and BaB has more ReLU to branch on. Therefore, Compensate($K$) is more accurate than $\beta$-CROWN, 
closing the gap with the upper bound from $20\%$ to $16.2\%$ ($8 \times 100$)
and from $17.6\%$ to $12\%$ ($8 \times 200$). This is very promising for the compensation idea, which could be used in many different ways than Compensate($K$) (see next Section).

More importantly, $\beta$-Crown, while efficient, often faces a (complexity) wall, when there number of branches becomes too extreme for BaB to work: either BaB succeeds fast, or it will not succeed even with very long time out (the reason why Refined BaB is more accurate on these networks). We verify that on a larger DNN learnt in a natural way, namely MLP $6 \times 500$, see Table \ref{tab:example3}.
We tested PureBaB with two time out, from the standard 30s to an extreme 2000s, to match the average run time of our adaptative implementation (fast mode). Only $3.5\%$ more of images could be ceritifed by extending the TO from 30s to 2000s, certifying $20\%$ less images than our prototype in the same time.
Notice that the refined BaB mode of $\beta$-Crown (using MILP for the first few layers) failed to run on this DNN (NoImplementation error code), and we were unable to fix the problem, which is not really surprising as running exact MILP with so many integer variables is not likely to be efficient, as 
confirmed while running PRIMA.
